### 2016-01-20 AssignmentIX/Assignment11-Ditz_Schroeder-WS15-16.tex
erste aufgabe des neuen blatts ist fertig

\newcommand{\points}[1]{
\newcounter{colmns}
\setcounter{colmns}{#1}
\stepcounter{colmns}
  \begin{vartab}{\thecolmns}
    \numbers{#1} & $\sum$ (6+2)\\\hline
    \emptyLine{\thecolmns}\\
  \end{vartab}
}

\begin{document}
%\header{Blatt}{Tutor}{Abgabedatum}{Vorlesung}{Bearbeiter}{Semester}{Anzahl Aufgaben}
\header{9}{Alexander Seitz}{18. January 2016}{Bioinformatics I}{\\Jonas Ditz \\\& Benjamin Schroeder}{WS 15/16}{4}

\section*{Theoretical Assignment - \textit{Overlap and Layout}}
\subsection*{graphs}
The graphs from the exercise are shown in Figure 1-3. Figure 1 is the overlap grap, Figure 2 is the minimal spanning tree and Figure 3 shows the layout. These figures are also publishable.

\begin{figure}[h]
	\includegraphics[width=\textwidth]{overlapGraph.png}
	\caption{Minimal-spanning-tree}
\end{figure}

\begin{figure}[h]
	\includegraphics[width=\textwidth]{minimal-Spanning-Tree.png}
	\caption{Minimal-spanning-tree}
\end{figure}

\begin{figure}[h]
	\includegraphics[width=\textwidth]{layout.png}
	\caption{layout}
\end{figure}

\subsection*{Is the Layout consitent?}
The overlaps are not consistent, because:
\begin{equation*}
o(3'f_6 - 3'f_8)=80
\end{equation*}
\section*{Theoretical Assignment - \textit{Overlap graph figure}}

\section*{Theoretical Assignment - \textit{Computing all overlaps using suffix tree}}

Finding all overlaps is basically a APSP (All-pairs-suffix-prefix) problem. Solving such a problem 
with a suffix tree approach has a time complexity of $O(n + k^2)$ where $n$ is the total length of 
all strings and $k$ is the count of strings. First all reads will be concatenated with a separation 
character between all reads and a terminal character at the end. Now it is rather simple to check, 
whether the prefix of one read matches with the suffix of another read. This can be done using the 
suffix tree since we know that each prefix starts with an separation character (one could put one 
separation character at the beginning of the string to make this statement true) and every suffix 
ends ether with the separation character or the terminal character. Since one has to go trough the 
whole concatenated string and compare the prefix of each read with the suffix of every other read, 
this approach has a total runtime complexity of $O(n+k^2)$.

\section*{Bonus - \textit{Challenges of your Project}}

One of the biggest challenges during our project was the framework we chose, Vaadin. We were not able 
to get it working of all computers we used. So in the end it was just one of us, who was able to work 
on the front-end layer. That slowed us down a lot. If we would start this project again from the scratch, 
Django would be a better choice. This framework comes with a development server and, therefore, would 
maybe make developing together easier. In the end we wasted a lot of time with Eclipse/Vaadin/Eclipse+Vaadin 
problems. Another challenge was the database schema for POODLE. We, especially Benjamin, talked a lot 
with people from Silke Wiesner's group at the MPI for Developmental Biology and that resulted in 
several schema changes during our project time.

>>>>

\newcommand{\points}[1]{
\newcounter{colmns}
\setcounter{colmns}{#1}
\stepcounter{colmns}
  \begin{vartab}{\thecolmns}
    \numbers{#1} & $\sum$ (10)\\\hline
    \emptyLine{\thecolmns}\\
  \end{vartab}
}

\begin{document}
%\header{Blatt}{Tutor}{Abgabedatum}{Vorlesung}{Bearbeiter}{Semester}{Anzahl Aufgaben}
\header{11}{Alexander Seitz}{18. January 2016}{Bioinformatics I}{\\Jonas Ditz \\\& Benjamin Schroeder}{WS 15/16}{2}

\section*{Theoretical Assignment - \textit{Assembly using de Bruijn graph}}

The de Bruijn graph for the reads provided on the assignment sheet would look like the graph in
Figure \ref{fig:4mere}, if one chooses $k = 4$. As one can see easily, there are not just one but 
four different Eulerian paths in the graph. Each path results in one of the following superstring:

\begin{align}
 S_1 &= AACCGTTAACGTAAACGT \nonumber \\
 S_2 &= AACCGTAAACGTTAACGT \nonumber \\
 S_3 &= AACCGTTAAACGTAACGT \nonumber \\
 S_4 &= AACCGTAACGTTAAACGT \nonumber
\end{align}

This happens due to the fact that there are nodes with more than one outgoing an incoming edge. If
one chooses $k = 5$ the resulting de Bruijn graph would look like the graph in Figure \ref{fig:5mere}.
It is obvious that the number of nodes is depended on $k$. With a bigger $k$ there are more nodes in 
the resulting de Bruijn graph. But in this case there is one problem. Since there are more than two 
unbalanced nodes in the de bruijn graph there is no Eulerian path.

\begin{figure}[h]
 \centering
 \includegraphics[width=0.7\textwidth]{deBruijnGraph_4mere.png}
 \caption{De Bruijn graph for the given reads and $k = 4$.}
 \label{fig:4mere}
\end{figure}

\begin{figure}[h]
 \centering
 \includegraphics[width=1\textwidth]{deBruijnGraph_5mere.png}
 \caption{De Bruijn graph for the given reads and $k = 5$.}
 \label{fig:5mere}
\end{figure}

\section*{Practical Assignment - \textit{Assemble the human mitochondrial genome}}
